### 2011-08-13 doc/theses/vitasek/diplomka.tex
Thesis: in progress

@@

TODO talk about all that are used - mention the dataset name, some characteristics (excel sheet)

\begin{table}
  \caption{List of realistic test data files}
  \label{table-experiments-data-realistic}
  \centering
  \begin{tabular}{l | r | c | c | l}
	Name  & Size [kb] & Vertices & Edges & Optimum \\
	\hline
	OVA1  & 4.5 & 29 & 43 & 0.45588235294117635 \\
	OVA2  & 11.9 & 23 & 36 & 0.1634615384615385  \\ 
	OVA3  & 237.6 & 31 & 47 & 0.25537156151635415 \\
	XMA-c & 1 807.7 & 1  & 0  & 0.7546666666666667  \\
	XMA-p & 13 748.3 & 1  & 0  & 0.2019306150568969  \\
	XMD   & 1 743.0 & 17 & 15 & 0.09786094165493507 \\
  \end{tabular}
\end{table}

TODO explain edges, vertices and say that opt is the known for alpha = beta = 1

@@

@@

TODO we have taken some of the text data and converted it to attributes like this, A B C.

TODO we can convert it back to keys!

\begin{table}
  \caption{List of realistic test data files with converted attributes}
  \label{table-experiments-data-converted}
  \centering
  \begin{tabular}{l | r | c | c | l}
	Name  & Size [kb] & Vertices & Edges & Optimum \\
	\hline
	MSH  & 3 100.5 & 1 & 0 & 0.5416472778036296 \\
	NTH  & 2 523.5 & 5 & 7 & 0.057918595422124436 \\ 
  \end{tabular}
\end{table}

@@

@@

TODO these are random graphs.

TODO talk about how they are created, about their structure.

TODO talk about some "density". Explain what is the max number of edges

TODO talk about the "size" test data

\begin{table}
  \caption{List of artificial test data files}
  \label{table-experiments-data-artificial}
  \centering
  \begin{tabular}{l | r | c | c | c | c | l}
	Name  & Size [kb] & Vertices & Edges & $|E| / |V|$ & $|E| / |E_{max}|$ & Optimum \\
	\hline
	100-100  & 8.4  & 99 & 95  & 0.95 & 0.02 & 0.836666666666667 \\
	100-200  & 13.0 & 96 & 174 & 1.81 & 0.04 & 0.726000000000000 \\ 
    100-1000 & 49.5 & 93 & 754 & 8.11 & 0.16 & 0.380952380952381 \\ 
  \end{tabular}
\end{table}

\begin{table}
  \caption{List of ``sized" artificial test data files}
  \label{table-experiments-data-artificial-size}
  \centering
  \begin{tabular}{l | r | c | c | c | c | l}
	Name  & Size [kb] & Vertices & Edges & $|E| / |V|$ & $|E| / |E_{max}|$ & Optimum \\
	\hline
	0-0     & 0.2  & 0  & 0   & -    & -    & 0.0                \\
	10-5    & 0.6  & 10 & 5   & 0.50 & 0.11 & 0.8500000000000002 \\ 
    20-20   & 1.7  & 18 & 13  & 0.72 & 0.08 & 0.7166666666666669 \\
   	30-45   & 3.1  & 29 & 43  & 1.48 & 0.11 & 0.7083333333333334 \\ 
	40-80   & 5.1  & 39 & 72  & 1.85 & 0.10 & 0.6950000000000002 \\
	50-125  & 7.5  & 48 & 111 & 2.31 & 0.10 & 0.6566666666666666 \\
	60-180  & 10.4 & 58 & 157 & 2.71 & 0.09 & 0.6214285714285716 \\
	70-245  & 13.8 & 67 & 205 & 3.06 & 0.09 & 0.5982142857142856 \\
	80-320  & 17.6 & 76 & 261 & 3.43 & 0.09 & 0.5791666666666667 \\
	90-405  & 21.9 & 86 & 352 & 4.09 & 0.10 & 0.528888888888889  \\
	100-500 & 26.7 & 91 & 388 & 4.26 & 0.09 & 0.4981818181818182 \\
  \end{tabular}
\end{table}

TODO explain why the optimum is getting smaller

@@

TODO why we have less than declared number? because of how we created it

>>>>

To understand these data sets, we will talk a little about heir origin, and a little more about their \textit{graph representation}. As mentioned earlier, % TODO link
the problem of finding the optimal ID set is in fact the problem of finding the maximum weighted independent set on a graph. Therefore, it might be interesting to actually see the graphs of these data sets and have some kind of numbers associated with them.

The former will be achieved with the help of GraphViz tool, % TODO link
where we will draw the graphs so that all the vertices represent the \textit{candidate AMs}, and the edges represent pairs of AMs that have nonempty intersection of their images (and thus cannot be in the same ID set together). Thus solving the maximal weighted IS on these graphs will be equivalent to solving our problem of optimal ID set.

The latter will come in form of tables containing information regarding the data sets, such their size, known optimum for $\alpha = \beta = 1$ and the numbers of vertices and edges in aforementioned graphs.

@@

From 3 different sources we collected 6 different data sets, called \jmodule{OVA1} - \jmodule{OVA3}, \jmodule{XMA-c}, \jmodule{XMA-p} and \jmodule{XMD}. Their summary is the Table \ref{table-experiments-data-realistic}, their graphs can be seen in Figure \ref{image-experiments-data-realistic}. Because the legal status of disclosing these data sets is unclear, we will refrain from identifying them beyond these artificial identifiers. Neither will they be included on the DVD distributed with this thesis.

TODO their DTD/XSD - do we have ID attributes?

\begin{table}
  \caption{List of realistic test data files}
  \bigskip
  \label{table-experiments-data-realistic}
  \centering
  \begin{tabular}{l | r | c | c | l}
	Name  & Size [kb] & $|V|$ & $|E|$ & Optimum \\
	\hline
	OVA1  & 4.5      & 29 & 43 & 0.45588235294117635 \\
	OVA2  & 11.9     & 23 & 36 & 0.1634615384615385  \\ 
	OVA3  & 237.6    & 31 & 47 & 0.25537156151635415 \\
	XMA-c & 1 807.7  & 1  & 0  & 0.7546666666666667  \\
	XMA-p & 13 748.3 & 1  & 0  & 0.2019306150568969  \\
	XMD   & 1 743.0  & 17 & 15 & 0.09786094165493507 \\
  \end{tabular}
\end{table}

@@

To interpret the data a little: \jmodule{OVA*} sets have quite interesting and perhaps a bit challenging graphs, but are not too big. We can consider them to be the ``typical" representants.

On the other hand, the \jmodule{XMA-*} sets are quite huge, but trivial: their only candidate AM will just get picked and that will be the end of the heuristic. We will see the performance of the other components of the whole system, dealing with huge data that are rather simple in the end.

Finally, the \jmodule{XMD} set is quite big and at the same time has non-trivial graph representation. We should see a performance more balanced between processing and finding the ID set here.

@@

We found 2 data sets to convert, \jmodule{MSH} and \jmodule{NTH}. Unfortunately, the same problem with disclosure as in the previous case applies here. None of these sets had any attributes before the conversion. Their summary is the Table \ref{table-experiments-data-converted}, their graphs can be seen in Figure \ref{image-experiments-data-converted}.

TODO schemas?

To address the conversion: in the case of \jmodule{MSH} we found 2 elements with values looking suspiciously like a key of the records contained in the file, and converted them to be attributes of these records using a simple XSL transformation. In the case of \jmodule{NTH} we converted all the values in sub-elements of the record elements to be the attributes of the records.

Why is this useful at all? Recall TODO link, where we said that ID attributes are a special case of XML keys. We can use this approach then to find XML keys: convert some suspicious data into attributes, find the optimal ID set and then create XML key based on this ID set.

\begin{table}
  \caption{List of realistic test data files with converted attributes}
  \bigskip
  \label{table-experiments-data-converted}
  \centering
  \begin{tabular}{l | r | c | c | l}
	Name  & Size [kb] & $|V|$ & $|E|$ & Optimum \\
	\hline
	MSH  & 3 100.5 & 1 & 0 & 0.5416472778036296 \\
	NTH  & 2 523.5 & 5 & 7 & 0.057918595422124436 \\ 
  \end{tabular}
\end{table}

@@

Again to do a little interpretation: in the case of \jmodule{MSH} we created 2 attributes, of which only one constituted a candidate AM. This is then the case similar to \jmodule{XMA-*} sets: quite large data, yet only one trivial ID attribute to be found.

In the case of \jmodule{NTH} we introduced more attributes, 8 to be precise. Out of them 5 proved to be candidate AMs, with 7 edges constraining them. This means we have quite a big set with relatively simple work to be done by the heuristics. This is about the change.

@@

As soon as we started experimenting with the data coming from the real world, it was obvious that they present no real challenge to our heuristics. After we build the model, we get the most complex graphs of some 31 vertices and 47 edges (see Table \ref{table-experiments-data-realistic}). We needed to create data that would really put pressure on our heuristics, put them to a real test. Our solutions is to approach from the other side: in the end, we will be solving the equivalent of IS problem on a graph created from XML data. Why not create the XML data to contain a more complex graph in the first place? What if we could specify how many vertices and edges we want, and get the according XML?

This is actually simple. Consider the following excerpt from an XML file:

\begin{scriptsize}
\begin{verbatim}
<graph>
  <vertex0 attr="-2968876296119015800"/>
  <vertex1 attr="1729745997570096518"/>
  <vertex2 attr="-9020549659620928934"/>
  ...
  <vertex99 attr="-7545982394508643394"/>

  <vertex82 attr="0"/><vertex21 attr="0"/>
  <vertex64 attr="1"/><vertex21 attr="1"/>
  <vertex44 attr="2"/><vertex2 attr="2"/>
  ...
  <vertex96 attr="99"/><vertex40 attr="99"/>
</graph>
\end{verbatim}
\end{scriptsize}

We want to create a graph with around $v$ vertices and $e$ edges. First, we introduce $v$ elements with names \texttt{vertex0} - \texttt{vertex\{V-1\}}. To constitute an AM, they need an attribute \texttt{attr}, but with large enough random values, so that these values won't conflict with any others. Second, for each of the $e$ edges we pick two \texttt{vertex*} elements at random, and give them the same value of their \texttt{attr}. This will ensure they cannot share the same ID set, thus effectively creating the edge in the graph representation.

The pseudocode for this is in listing \ref{listing-random-data}.

\begin{algorithm}
\caption{Random XML data creation}
\label{listing-random-data}
\begin{algorithmic}
\REQUIRE $v$ requested number of vertices
\REQUIRE $e$ requested number of edges
\ENSURE XML file content
\PRINT \texttt{<graph>}
\FOR{$i = 1 \to |V|$}
	\STATE $R \gets RANDOM$
	\PRINT \texttt{<vertex\textit{i} attr="\textit{R}">}
\ENDFOR
\FOR{$i = 1 \to |E|$}
	\STATE $v1 \gets RANDOM(|V|)$
	\STATE $v2 \gets RANDOM(|V|)$
	\PRINT \texttt{<vertex\textit{v1} attr="\textit{i}"> <vertex\textit{v2} attr="\textit{i}">}
\ENDFOR
\PRINT \texttt{</graph>}
\RETURN
\end{algorithmic}
\end{algorithm}

With this process in place we can create as much data as we wish with any combination of $v$ and $e$ we want.

There is one interesting characteristic that can describe random graphs like this, and that is the \textit{density}. This can be defined in various ways, we will use two different interpretations for a while. The first is $\frac{|E|}{|V|}$, that is, how many edges are there for one vertex (multiplied by 2 we'd get the average degree of the vertices). The second, perhaps more interesting is $\frac{|E|}{E_{max}}$, where $E_{max} = \frac{|V|.(|V|-1)}{2}$. This is the density as the ratio of edges that are to all edges that could be in a complete graph with $|V|$ vertices.

We have created 3 sets to use in experiments along with the realistic and converted sets, these are called \jmodule{100-100}, \jmodule{100-200} and \jmodule{100-1000}. Note that the name is always in the form $v-e$.

Also, we will need data of comparably similar characteristics but varying size, to study the effects of size on the run times of experiments. For this reason we created 11 more sets, from \jmodule{0-0} as the trivial one to \jmodule{100-500} as the biggest one. We wanted to keep density the same among these sets, and we picked the $\frac{|E|}{E_{max}}$ density interpretation for this.

TODO this is on the DVD.

The summary is the Table \ref{table-experiments-data-artificial} and Table \ref{table-experiments-data-artificial-size}, these tables contain 2 new columns: values of density in both interpretations we introduced. Some graphs can be seen in Figure \ref{image-experiments-data-artificial}.

While studying the tables you might notice that the actual numbers $|V|$ and $|E|$ do not match to the $v$ and $e$ in the names of the sets. This is because how the random generation algorithm works: it might pick the same edge twice, which will automatically render it unsuitable for the ID set (proof is homework). % TODO too much?:)
Because of the so-called Birthday paradox, % TODO link
this will happen more with higher $e$.

\begin{table}
  \caption{List of artificial test data files}
  \bigskip
  \label{table-experiments-data-artificial}
  \centering
  \begin{tabular}{l | r | c | c | c | c | l}
	Name  & Size [kb] & $|V|$ & $|E|$ & $\frac{|E|}{|V|}$ & $\frac{|E|}{E_{max}}$ & Optimum \\
	\hline
	100-100  & 8.4  & 99 & 95  & 0.95 & 0.02 & 0.836666666666667 \\
	100-200  & 13.0 & 96 & 174 & 1.81 & 0.04 & 0.726000000000000 \\ 
    100-1000 & 49.5 & 93 & 754 & 8.11 & 0.16 & 0.380952380952381 \\ 
  \end{tabular}
\end{table}

\begin{table}
  \caption{List of ``sized" artificial test data files}
  \bigskip
  \label{table-experiments-data-artificial-size}
  \centering
  \begin{tabular}{l | r | c | c | c | c | l}
	Name  & Size [kb] & $|V|$ & $|E|$ & $\frac{|E|}{|V|}$ & $\frac{|E|}{E_{max}}$ & Optimum \\
	\hline
	0-0     & 0.2  & 0  & 0   & -    & -    & 0.0                \\
	10-5    & 0.6  & 10 & 5   & 0.50 & 0.11 & 0.8500000000000002 \\ 
    20-20   & 1.7  & 18 & 13  & 0.72 & 0.08 & 0.7166666666666669 \\
   	30-45   & 3.1  & 29 & 43  & 1.48 & 0.11 & 0.7083333333333334 \\ 
	40-80   & 5.1  & 39 & 72  & 1.85 & 0.10 & 0.6950000000000002 \\
	50-125  & 7.5  & 48 & 111 & 2.31 & 0.10 & 0.6566666666666666 \\
	60-180  & 10.4 & 58 & 157 & 2.71 & 0.09 & 0.6214285714285716 \\
	70-245  & 13.8 & 67 & 205 & 3.06 & 0.09 & 0.5982142857142856 \\
	80-320  & 17.6 & 76 & 261 & 3.43 & 0.09 & 0.5791666666666667 \\
	90-405  & 21.9 & 86 & 352 & 4.09 & 0.10 & 0.528888888888889  \\
	100-500 & 26.7 & 91 & 388 & 4.26 & 0.09 & 0.4981818181818182 \\
  \end{tabular}
\end{table}

@@

To interpret Tables \ref{table-experiments-data-artificial} and \ref{table-experiments-data-artificial-size}: we get 3 sets of different sizes and densities in the first one, which should work very nice with our heuristics. The $|V|$ and $|E|$ numbers are orders of magnitude higher than with any realistic (or converted) data set we are using, and this will put real stress on the heuristics, as we shall see in the next section.

In the second table we aimed for the $\frac{|E|}{E_{max}}$ density of $0.1 = 10\%$, and we can see that this was quite nicely achieved. There is an interesting observation to be made here: the optimum is steadily decreasing with the increasing overall graph size. This intuitively suggests that the maximal quality theoretically achievable has to do with the $\frac{|E|}{|V|}$ density, not with the one we fixed. Exploration of this phenomenon is among the possibilities of future work of this thesis.
